%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass[masters]{ucbthesis}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage[normalem]{ulem}
\usepackage{soul}
\usepackage{array}
\usepackage{amssymb}
\usepackage{extarrows}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{framed}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber,
natbib=true,
style=authoryear,
sorting=none,
isbn=true,
doi=true,
url=true,
]{biblatex}

\makeatletter
% special thanks to this guy
\newrobustcmd*{\parentexttrack}[1]{%
  \begingroup
  \blx@blxinit
  \blx@setsfcodes
  \blx@bibopenparen#1\blx@bibcloseparen
  \endgroup}

\AtEveryCite{%
  \let\parentext=\parentexttrack%
  \let\bibopenparen=\bibopenbracket%
  \let\bibcloseparen=\bibclosebracket}

\makeatother
\addbibresource{library.bib}

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{wrapfig}
\usepackage{wasysym}
\usepackage{enumitem}
\usepackage{adjustbox}
\usepackage{ragged2e}
\usepackage[svgnames,table]{xcolor}
\usepackage{tikz}
% csv files and plots
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{csvsimple-l3}
\usepackage{longtable}
\usepackage{changepage}
\usepackage{setspace}
\usepackage{hhline}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{float}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{fancyhdr}
\usetikzlibrary{shapes.symbols,shapes.geometric,shadows,arrows.meta,
positioning,fit,backgrounds}
\tikzset{>={Latex[width=1.5mm,length=2mm]}}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage{listings}

\usepackage{listings}
\lstset{
    language=C++, % Set the programming language
    backgroundcolor=\color{black!5}, % Light grey background
    basicstyle=\footnotesize\ttfamily, % Font style and size
    keywordstyle=\color{blue}, % Keyword color
    commentstyle=\color{green!60!black}, % Comment color
    stringstyle=\color{red}, % String color
    showstringspaces=false,
    numbers=left, % Line numbers on the left
    numberstyle=\tiny\color{gray}, % Line number style
    frame=single, % Add a frame around the code
    breaklines=true, % Automatic line breaking
    captionpos=b, % Caption position at the bottom
}


\urlstyle{same}

\renewcommand{\_}{\kern-1.5pt\textunderscore\kern-1.5pt}

%https://tex.stackexchange.com/questions/116534/lstlisting-line-wrapping
% I copied code from here to format things properly
% Credit to: Henri Menke who answered on StackExchange
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

%%  Import package so that we can split the tex files into subfolders
% for easier compile and real-time edit

\usepackage{import}

%%%%%%%%%%%%%% for transfer function controllers %%%%%%%%%%%%%%
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
%%%% set citation style to square brackets


\begin{document}

% Declarations for Front Matter

\title{Derivations for Tampines Steam Tables and TAMPINES}
\author{Tom Lifesaver}
%\degreesemester{Spring}
%\degreeyear{1995}
%\degree{Doctor of Philosophy}
%\chair{Professor Richard Francis Sony}
%\othermembers{Professor Roger Spam \\
%  Associate Professor Michael Chex}
%% For a co-chair who is subordinate to the \chair listed above
%% \cochair{Professor Benedict Francis Pope}
%% For two co-chairs of equal standing (do not use \chair with this one)
%% \cochairs{Professor Richard Francis Sony}{Professor Benedict Francis Pope}
%\numberofmembers{3}
%% Previous degrees are no longer to be listed on the title page.
%% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
%\field{Mathematics}
%% Designated Emphasis -- this is optional, and rare
%% \emphasis{Colloidal Telemetry}
%% This is optional, and rare
%% \jointinstitution{University of Western Maryland}
%% This is optional (default is Berkeley)
%% \campus{Berkeley}
%
%% For a masters thesis, replace the above \documentclass line with
%% \documentclass[masters]{ucbthesis}
%% This affects the title and approval pages, which by default calls this
%% document a "dissertation", not a "thesis".
%
%\maketitle
%% Delete (or comment out) the \approvalpage line for the final version.
%\approvalpage
%\copyrightpage

\include{abstract}

\begin{frontmatter}


% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\clearpage
\listoftables

\begin{acknowledgements}
\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\part{Introduction}

\part{What is TAMPINES}

\part{Steam Tables}

\part{Rankine Cycle Derivations}

In this part, we consider a four control volume system representing a 
simple Rankine cycle. This will form the theoretical basis to 

\chapter{Mass, Energy and Momentum Balance}

In OpenFOAM, the PIMPLE algorithm is often used to solve for incompressible 
fluid flow \citep{devolder2015review}, this is one of the many standard 
solvers for solving flow in OpenFOAM \citep{penttinen2011pimplefoam}.
Jasak has published this in his PhD thesis, available online,
\citep{jasak1996error} on page 143-149.

\section{Raw forms of mass, momentum and energy balance}

Let's consider mass balance and momentum balance:
\begin{equation}
	\frac{\partial m}{\partial t} = \sum_i m_i
\end{equation}


\begin{equation}
	\frac{\partial (\rho \textbf{u})}{\partial t} 
	+ \nabla \cdot (\rho \textbf{u} \textbf{u})
	= - \nabla P
	+ \nabla \cdot \tau 
	+ \rho g
\end{equation}

$\nabla \cdot \tau$ represents shear, $\rho g$ represents the gravity 
body force, $- \nabla P$ represents the pressure gradient.

\begin{equation}
	V_j \frac{\partial \rho_i h_i}{\partial t} + \sum_j \dot{m}_j h_j
	= Q_i - W_i 
	+ \sum_j \mathbf{\tau}_j \cdot \mathbf{u}_j
\end{equation}

Where $+ \sum_j \mathbf{\tau}_j \cdot \mathbf{u}_j $ is basically the 
friction loss terms where energy is transformed from kinetic energy 
to sensible heat. Q and W represent the thermodynamic work

\section{converting forms to mass flowrate}

It is often more convenient to convert the momentum equations to be 
in terms of mass flowrates. This is for 1D case.

To do so, we transform this by consider this 3D equation.
\begin{equation}
	\frac{\partial (\rho \textbf{u})}{\partial t} 
	+ \nabla \cdot (\rho \textbf{u} \textbf{u})
	= - \nabla P
	+ \nabla \cdot \tau 
	+ \rho g
\end{equation}

We need to consider just the x component, for a simple 1D case.

Then multiply by the cross sectional area.

\section{Approach to Solving nonlinear Equations}


Jasak in his PhD thesis outlines the methodology for solving this set of 
equation \citep{jasak1996error} on page 143-149. These were the foundations 
for what would eventually become OpenFOAM \cite{jasak2007openfoam} licensed 
under GPLv3.

The momentum equation is quadratic in u, which makes the equation solution 
nonlinear. Solving this nonlinear equation would either need a nonlinear 
solver such as Newton Raphson method, or solving a linearised form of 
these equations, and iterating to convergence. The latter was chosen 
\citep{jasak1996error}.

\begin{equation}
	\nabla \bullet (\mathbf{u u}) 
	= \sum_f \mathbf{S} \cdot \mathbf{u}_f\mathbf{u}_f
	= \sum_f F \mathbf{u}_f
	= a_p \mathbf{u}_p + \sum_N a_N \textbf{u}_n
\end{equation}

\subsection{Finite Volume Discretisaion, Face and Centroid Properties}
Section 3.3.1 within Jasak's thesis explains these terms.

First, we need to understand that Gauss's theorem is used in Finite 
Volume analysis, which is the bedrock of OpenFOAM solvers. 
$\sum_f \mathbf{S} \cdot \mathbf{u}_f\mathbf{u}_f$ represent the surface 
vector $\mathbf{S}$ and this will have a dot product with the velocity 
at the faces of the control volume $\mathbf{u}_f$.

Note that the subscript f represents values at the faces.

$\sum_f \mathbf{S} \cdot \mathbf{u}_f\mathbf{u}_f$
can be written in shorthand $\sum_f F \mathbf{u}_f$.

Now, we may not have the values of the velocities or fluxes at the 
faces. So we have to estimate it based on the properties at the 
centroid of the cell denoted $p$, and those of the neighbours denoted $n$.

Thus, the surface values denoted $f$ are estimated based on the values 
at the cell centroids $p$ and $n$ based on some kind of interpolation 
scheme. This can be linear interpolation, or some upwind differencing scheme.
This is described in page 81 of Jasak's thesis \citep{jasak1996error}.

With linear interpolation at the face, the solution can become unbounded.
With upwind differencing, the solution is bounded, but diffusive. Much 
of the discussion revolves around how to estimate properties at the 
faces. This is why a whole host of differencing schemes comes up. Eg. 
at page 110 of Jasak's thesis \citep{jasak1996error}.

\subsection{Linearisation and Predictor values}

Given a very nonlinear system of equations. It was desirable to 
consider solving a linearised version of the momentum equations.
We want to solve for the velocity $\mathbf{u}^{t + \Delta t}$ at the 
next timestep. But we only know $\mathbf{u}^{t }$.

We know that the face fluxes should be based on values in the cell 
centroids for both the cell $p$ and that of the neighbouring cell $n$,
hence we use:

$$a_p \mathbf{u}_p + \sum_N a_N \textbf{u}_n$$

However, we have $\mathbf{u u}$ terms. To solve this problem and linearise 
this, we can use $\mathbf{u}^t \mathbf{u}^{t + \Delta t}$. Basically, 
$\mathbf{u}^t$ is known. This is the velocity at the previous timestep,
and is therefore a constant. This will then help form the respective 
$a_p$ and $a_n$ constants so that the equations are linear.

Let's go back to our equations:
\begin{equation}
	\frac{\partial (\rho \textbf{u})}{\partial t} 
	+ \nabla \cdot (\rho \textbf{u} \textbf{u})
	= - \nabla P
	+ \nabla \cdot \tau 
	+ \rho g
\end{equation}

In the incompressible assumption,

\begin{equation}
	\frac{\partial (\textbf{u})}{\partial t} 
	+ \nabla \cdot (\textbf{u} \textbf{u})
	= - \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

We can do time discretisation:

\begin{equation}
	\frac{\textbf{u}^{t+ \Delta t} - \textbf{u}^t}{\Delta t} 
	+ \nabla \cdot (\textbf{u} \textbf{u})
	= - \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

Now, remember, based on divergence theorem we need fluxes or properties 
at the faces:

\begin{equation}
	\frac{\textbf{u}^{t+ \Delta t} - \textbf{u}^t}{\Delta t} 
	+ \sum_f \mathbf{S} \cdot \mathbf{u}_f \mathbf{u}_f
	= - \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

And since we cannot find properties at the faces, we have to estimate 
them using properties at the centroids.

Just to be clear, let's label the velocities at their proper locations,
and then linearise the equation using some semi implicit time 
discretisation:

\begin{equation}
	\frac{\textbf{u}^{t+ \Delta t}_p - \textbf{u}^t_p}{\Delta t} 
	+ \sum_f \mathbf{S} \cdot \mathbf{u}_f^t \mathbf{u}_f^{t + \Delta t} 
	= - \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

The fluxes at the faces can be further estimated using some interpolation 
scheme. The easiest to understand is a linear scheme. Herein, we use some 
weighting factor to estimate the flux at the face between the cell 
and the other cell.

$$ \textbf{u}_f^t = w \textbf{u}_p^t + (1-w) \textbf{u}_{n}^t  $$

Substituting this back:

\begin{equation}
	\frac{\textbf{u}^{t+ \Delta t}_p - \textbf{u}^t_p}{\Delta t} 
	+ \sum_f \mathbf{S}\cdot (w \textbf{u}_p^t + (1-w) \textbf{u}_{n}^t) 
	(w \textbf{u}_p^{t + \Delta t} + (1-w) \textbf{u}_{n}^{t + \Delta t})
	= - \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

To make life easier for us, we can choose not to even bother with 
explicitly solving for the velocities of the neigbours

\begin{equation}
	\frac{\textbf{u}^{t+ \Delta t}_p - \textbf{u}^t_p}{\Delta t} 
	+ \sum_f \mathbf{S}\cdot ( w \textbf{u}_p^t + (1-w) \textbf{u}_{n}^t) 
	(w \textbf{u}_p^{t + \Delta t} + (1-w) \textbf{u}_{n}^{t })
	= 
	- \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

Messy as the equation is, we now have a linear equation in velocities 
for each centroid at the next timestep $u^{t + \Delta t}_p$ 
and $u^{t + \Delta t}_n$.

We can then simplify this equation as such:


\begin{equation}
	a_p \textbf{u}_p^{t + \Delta t} = \textbf{H}(\textbf{u}^t)
	- \nabla P/\rho
	+ \nabla \cdot \tau /\rho
	+ g
\end{equation}

\begin{equation}
	\textbf{u}_p^{t + \Delta t} = \frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\end{equation}


Of course, the estimates for $\mathbf{u}^{t + \Delta t}$ are not going to 
be accurate. They are initial guess or \textbf{predicted} values.

We need to do some further correction to these predicted values in order 
to arrive at the correct answer.

\subsection{Corrector steps}

So far, we've only bothered with the momentum equation and ignored 
mass balance (or continuity in this case). Thus, there is 
no guarantee that the mass balance is fulfilled.

The mass balance is:

$$\nabla \cdot \textbf{u}_p = 0$$

We then apply Gauss's divergence theorem to obtain:

$$\nabla \cdot \textbf{u}_p = \sum_f \textbf{S} \cdot \textbf{u}_f = 0$$

Now again, how do we deal with the face fluxes again?
We need to interpolate.

Suppose again that we can use some weighting factor (linear 
interpolation scheme), that on a face between a point and its neighbouring 
cell:

$$ u_f = w u_p + (1-w) u_n $$


\begin{equation*}
	\textbf{u}_p^{t + \Delta t} = \frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\end{equation*}

since

\begin{equation*}
	\textbf{u}_p^{t + \Delta t} = 0
\end{equation*}

Then:


$$\nabla \cdot 
\left( 
	\frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\right)
=  0$$

Moving the pressure term to the right:

$$\nabla \cdot 
\left( 
	\frac{\textbf{H}(\textbf{u}^t)}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\right)
=  \nabla \cdot \frac{P/\rho}{a_p}
$$


This is equation 3.141 in page 146 of Jasak's thesis \citep{jasak1996error}.
Of course, $\tau$ and $g$ were not included.

Now, if we apply divergence theorem on the H terms:

$$\nabla \cdot 
\left( 
	\frac{\textbf{H}(\textbf{u}^t)}{a_p}
\right)
=  
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{\textbf{H}(\textbf{u}^t)}{a_p}
\right)_f
$$

We also need to discretise the pressure side

$$
\nabla \cdot \frac{P/\rho}{a_p}
= 
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
$$

With this discretisation, we obtain

$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{\textbf{H}(\textbf{u}^t)}{a_p}
\right)_f
= 
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
$$

With the interpolation:
$$ u_f = w u_p + (1-w) u_n $$

the matrices can be set up.
This is equation 3.143 in Jasak's thesis \citep{jasak1996error}.

So basically, the continuity equation is combined with the momentum 
equation to give a combined corrector equation.

\section{PISO Algorithm}

Pressure Implicit Splitting of Operators (PISO).

This includes a few steps, outlined in Jasak's thesis \citep{jasak1996error}:

\subsection{Step 1}

This is the momentum predictor step, we can predict a velocity field 
using the previous steps
\begin{equation}
	\textbf{u}_p^{predictor} = \frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\end{equation}

\subsection{Step 2}

With this, we can formulate a new $\frac{\textbf{H}(\textbf{u}^t)}{a_p}$:

\begin{equation}
	\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
\end{equation}

With this, we can substitute into the mass balance (pressure equation):
$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
=
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
\right)_f
$$

This allows for a new pressure field. Note that the gravity terms are missing 
because hydrostatic pressure is ignored.

$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
$$

It is conceivable that this results in some small matrix where 
pressure fields are simultaneously solved.

This is the pressure solution within Jasak's thesis page 147 \citep{jasak1996error}.

\subsection{Step 3}

We now have a new pressure field based on mass balance:

$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
$$

And a new predicted H field based on the initial predictor step:

$$
\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
$$

Now, the predictor H field is not guaranteed to allow continuity with 
the old velocity, but since it was used to generate a new pressure 
field using the continuity equation, it should be able to guarantee 
creation of a new velocity field which will satisfy continuity equation.


\begin{equation}
	\textbf{u}_p^{corrected} = \frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
	- \left(\frac{\nabla P/\rho}{a_p}\right)_\text{corrected pressure field}
\end{equation}

\subsection{repeated iterations}
This forms the basis for iterating through both mass and momentum balances.
This is one round of PISO prediction-correction. 

We may iterate this several times until the residual is small between 
consecutive iterations. 

Now, it must be noted that during PISO, we do NOT repeat step 1 
to recalculate as mentioned in page 148 Jasak's thesis \cite{jasak1996error}:

\begin{equation*}
	\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
\end{equation*}

The H term is kept constant through the whole correction sequence and 
changed in only the next momentum predictor \citep{jasak1996error}.





\section{SIMPLE Algorithm}

The SIMPLE algorithm is similar, except that it is solving for steady state 
solutions, and uses large time step. 

The key difference is rather than skip Step 1 within loops, we repeat 
step 1 for momentum predictor. This is the key difference between 
PISO and SIMPLE. It is that we run the momentum predictor step every 
time in SIMPLE. In this regard, it is more important to re-calculate 
H in every iteration so that non-linearities are accounted for.

Hence, between iterations, there is no need to fully 
resolve pressure-velocity coupling. Which means 
between iterations, one does not need to actively ensure that they are 
solved simultaneously. 


\subsection{Step 1}

Same as PISO,
this is the momentum predictor step, we can predict a velocity field 
using the previous steps
\begin{equation}
	\textbf{u}_p^{predictor} = \frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\end{equation}

However, before going to step 2, we 

\subsection{Step 2}

Also same as PISO, we do the pressure equation step

With this, we can formulate a new $\frac{\textbf{H}(\textbf{u}^t)}{a_p}$:

\begin{equation}
	\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
\end{equation}

With this, we can substitute into the mass balance (pressure equation):
$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
=
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
\right)_f
$$

This allows for a new pressure field. Note that the gravity terms are missing 
because hydrostatic pressure is ignored.

$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
$$

It is conceivable that this results in some small matrix where 
pressure fields are simultaneously solved.

This is the pressure solution within Jasak's thesis page 
147 \citep{jasak1996error}.

\subsection{Step 3: Momentum Corrector}


We now have a new pressure field based on mass balance:

$$
\sum_f \textbf{S}_f \cdot
\left( 
	\frac{P/\rho}{a_p}
\right)_f
$$

And a new predicted H field based on the initial predictor step:

$$
\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
$$

Now, the predictor H field is not guaranteed to allow continuity with 
the old velocity, but since it was used to generate a new pressure 
field using the continuity equation, it should be able to guarantee 
creation of a new velocity field which will satisfy continuity equation.


\begin{equation}
	\textbf{u}_p^{corrected} = \frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
	- \left(\frac{\nabla P/\rho}{a_p}\right)_\text{corrected pressure field}
\end{equation}

\subsection{repeated iterations}
This forms the basis for iterating through both mass and momentum balances.
This is one round of SIMPLE prediction-correction. 

We may iterate this several times until the residual is small between 
consecutive iterations. 

Now, it must be noted that previously during PISO, we do NOT repeat step 1 
to recalculate as mentioned in page 148 Jasak's thesis \cite{jasak1996error}:

\begin{equation*}
	\frac{\textbf{H}(\textbf{u}^{predictor})}{a_p}
\end{equation*}

Now, in SIMPLE, we do change H during subsequent iterations in the 
momentum predictor step \citep{jasak1996error}.

\href{https://www.youtube.com/watch?v=ahdW5TKacok}{
	Dr Aiden's YouTube videos were helpful in understanding this regard 
	too.
}

\subsection{Relaxation Factors}

However, between each iteration, timesteps may be large. Hence, the 
solution may experience some form of instability.
Therefore, there is the introduction of relaxation factors which updates 
the velocity slowly between successive momentum predictor steps.

Jasak mentions the use of velocity under-relaxation factors $\alpha_u$.
as well as pressure under-relaxation factors $\alpha_p$.

He outlined the following steps on page 148-149 of his thesis 
\citep{jasak1996error}. For pressure:

$$p^{new} = p^{old} + \alpha_p (p^{predictor} - p^{old})$$

For momentum, a similar thing applies. Though for momentum $\alpha_u$
is usually around 0.8 and for pressure $\alpha_p$ is around 0.2. Jasak 
writes this in page 149 of his thesis \citep{jasak1996error}.

\section{PIMPLE Algorithm}

For the PIMPLE algorithm, this is a merger between PISO and SIMPLE 
algorithms \citep{greenshields2022notes}.

Basically, every SIMPLE iteration now has the option to perform a set 
number of PISO iterations as well. That is perform the corrector steps 
multiple times per SIMPLE iteration. The PISO loops are known as inner loops,
whereas the outer loops are known as SIMPLE loops.

It is then up to the user to know how many loops are required.

With this understanding, it is good to try and apply this thought process 
to our four control volume situation.

\section{Application to Rankine Cycle}


\subsection{Mass, Energy and Momentum Balance Equations}
Consider four control volumes. We consider the following PDEs:


\begin{equation}
	\frac{\partial m}{\partial t} = \sum_i m_i
\end{equation}


\begin{equation}
	\frac{\partial (\rho \textbf{u})}{\partial t} 
	+ \nabla \cdot (\rho \textbf{u} \textbf{u})
	= - \nabla P
	+ \nabla \cdot \tau 
	+ \rho g
\end{equation}

$\nabla \cdot \tau$ represents shear, $\rho g$ represents the gravity 
body force, $- \nabla P$ represents the pressure gradient.

\begin{equation}
	V_j \frac{\partial \rho_i h_i}{\partial t} + \sum_j \dot{m}_j h_j
	= Q_i - W_i 
	+ \sum_j \mathbf{\tau}_j \cdot \mathbf{u}_j
\end{equation}

For a 1D case, we can consider Equations from Modelica. This was 
the excerpt from my dissertation \citep{ong2024digital}:
\begin{framed}
	
The mass balance equation for a 1D geometry is \citep{Casella2003}.

\begin{equation} \label{eqn:modelica-mass-balance-1D}
	A_{XS} \frac{\partial \rho}{\partial t} + \frac{\partial \dot{m}}{\partial x} = 0
\end{equation}

Where $A_{XS}$ is the cross sectional area, $\rho$ is the density of 
the differential volume $dV$, $\dot{m}$ is the mass flow rate, and 
$x$ is the length coordinate in the direction of flow.

The momentum balance is \citep{Casella2003}:

\begin{equation} \label{eqn:modelica-momentum-balance-1D}
	\frac{\partial \dot{m}}{\partial t} 
	+ \frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation}

In Equation~\ref{eqn:modelica-momentum-balance-1D}, in addition to the 
terms already defined for Equation~\ref{eqn:modelica-mass-balance-1D},
we have Fanning friction factor $f_{fanning}$, wetted perimeter $P_w$, 
acceleration due to gravity $g$, height for calculating hydrostatic pressure 
$z$ and pressure $P$. The absolute mass flow rate $\dot{m}$ is used to 
account for flow reversal in the momentum equation. Additionally 
the derivative $\frac{\partial \dot{m}}{\partial t}$ is often set to zero 
to avoid fast pressure oscillations \citep{Casella2003}. The timescale of 
these oscillations is quite often connected to the speed of sound in that 
medium. 

The energy balance is \citep{Casella2003}:

\begin{equation} \label{eqn:modelica-energy-balance-1D}
	\rho A_{XS} \frac{\partial h_{enthalpy}}{\partial t} 
	+ A_{XS} \frac{\partial P}{\partial t} = 
	- \rho A_{XS} u \frac{\partial h_{enthalpy}}{\partial x}
	+ q_{wall} \frac{A_{wall}}{L}
\end{equation}

In Equation~\ref{eqn:modelica-energy-balance-1D}, in addition to the previously 
defined terms for Equation~\ref{eqn:modelica-mass-balance-1D} and 
\ref{eqn:modelica-momentum-balance-1D}, I define $h_{enthalpy}$
for specific enthalpy per unit mass for the control volume, average 
fluid velocity $u$ and $q_{wall}$, the heat flux received or lost through 
lateral surfaces. $h_{enthalpy}$ is denoted as such to prevent confusion 
with heat transfer coefficient, also commonly denoted $h$. 
$\frac{A_{wall}}{L}$ is the heated surface area of the wall $A_{wall}$ 
per unit length, where $L$ is the length of the 1D geometry. For tubes,
$\frac{A_{wall}}{L} = P_w$ \citep{Casella2003}.
\end{framed}

If we consider the 1D balances:

\begin{equation*} 
	A_{XS} \frac{\partial \rho}{\partial t} + \frac{\partial \dot{m}}{\partial x} = 0
\end{equation*}

\begin{equation*} 
	\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

\begin{equation*} 
	\rho A_{XS} \frac{\partial h_{enthalpy}}{\partial t} 
	+ A_{XS} \frac{\partial P}{\partial t} = 
	- \rho A_{XS} u \frac{\partial h_{enthalpy}}{\partial x}
	+ q_{wall} \frac{A_{wall}}{L}
\end{equation*}

Of course, for the equation of state, we can use a real gas equation 
or steam table. That is done in tampines-steam-tables.

\subsection{rhoPimpleFoam}

For the algorithm considered, we are going to translate from
rhoPimpleFoam to what we want:

\begin{lstlisting}[caption=rhoPimpleFoam.C, label={lst:rhoPimpleFoam}]
/*---------------------------------------------------------------------------*\
  =========                 |
  \\      /  F ield         | OpenFOAM: The Open Source CFD Toolbox
   \\    /   O peration     |
    \\  /    A nd           | www.openfoam.com
     \\/     M anipulation  |
-------------------------------------------------------------------------------
    Copyright (C) 2011-2017 OpenFOAM Foundation
    Copyright (C) 2019 OpenCFD Ltd.
-------------------------------------------------------------------------------
License
    This file is part of OpenFOAM.

    OpenFOAM is free software: you can redistribute it and/or modify it
    under the terms of the GNU General Public License as published by
    the Free Software Foundation, either version 3 of the License, or
    (at your option) any later version.

    OpenFOAM is distributed in the hope that it will be useful, but WITHOUT
    ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
    FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License
    for more details.

    You should have received a copy of the GNU General Public License
    along with OpenFOAM.  If not, see <http://www.gnu.org/licenses/>.

Application
    rhoPimpleFoam

Group
    grpCompressibleSolvers

Description
    Transient solver for turbulent flow of compressible fluids for HVAC and
    similar applications, with optional mesh motion and mesh topology changes.

    Uses the flexible PIMPLE (PISO-SIMPLE) solution for time-resolved and
    pseudo-transient simulations.

Note
   The motion frequency of this solver can be influenced by the presence
   of "updateControl" and "updateInterval" in the dynamicMeshDict.

\*---------------------------------------------------------------------------*/

#include "fvCFD.H"
#include "dynamicFvMesh.H"
#include "fluidThermo.H"
#include "turbulentFluidThermoModel.H"
#include "bound.H"
#include "pimpleControl.H"
#include "pressureControl.H"
#include "CorrectPhi.H"
#include "fvOptions.H"
#include "localEulerDdtScheme.H"
#include "fvcSmooth.H"

// * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * //

int main(int argc, char *argv[])
{
    argList::addNote
    (
        "Transient solver for compressible turbulent flow.\n"
        "With optional mesh motion and mesh topology changes."
    );

    #include "postProcess.H"

    #include "addCheckCaseOptions.H"
    #include "setRootCaseLists.H"
    #include "createTime.H"
    #include "createDynamicFvMesh.H"
    #include "createDyMControls.H"
    #include "initContinuityErrs.H"
    #include "createFields.H"
    #include "createFieldRefs.H"
    #include "createRhoUfIfPresent.H"

    turbulence->validate();

    if (!LTS)
    {
        #include "compressibleCourantNo.H"
        #include "setInitialDeltaT.H"
    }

    // * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * //

    Info<< "\nStarting time loop\n" << endl;

    while (runTime.run())
    {
        #include "readDyMControls.H"

        // Store divrhoU from the previous mesh so that it can be mapped
        // and used in correctPhi to ensure the corrected phi has the
        // same divergence
        autoPtr<volScalarField> divrhoU;
        if (correctPhi)
        {
            divrhoU.reset
            (
                new volScalarField
                (
                    "divrhoU",
                    fvc::div(fvc::absolute(phi, rho, U))
                )
            );
        }

        if (LTS)
        {
            #include "setRDeltaT.H"
        }
        else
        {
            #include "compressibleCourantNo.H"
            #include "setDeltaT.H"
        }

        ++runTime;

        Info<< "Time = " << runTime.timeName() << nl << endl;

        // --- Pressure-velocity PIMPLE corrector loop
        while (pimple.loop())
        {
            if (pimple.firstIter() || moveMeshOuterCorrectors)
            {
                // Store momentum to set rhoUf for introduced faces.
                autoPtr<volVectorField> rhoU;
                if (rhoUf.valid())
                {
                    rhoU.reset(new volVectorField("rhoU", rho*U));
                }

                // Do any mesh changes
                mesh.controlledUpdate();

                if (mesh.changing())
                {
                    MRF.update();

                    if (correctPhi)
                    {
                        // Calculate absolute flux
                        // from the mapped surface velocity
                        phi = mesh.Sf() & rhoUf();

                        #include "correctPhi.H"

                        // Make the fluxes relative to the mesh-motion
                        fvc::makeRelative(phi, rho, U);
                    }

                    if (checkMeshCourantNo)
                    {
                        #include "meshCourantNo.H"
                    }
                }
            }

            if (pimple.firstIter() && !pimple.SIMPLErho())
            {
                #include "rhoEqn.H"
            }

            #include "UEqn.H"
            #include "EEqn.H"

            // --- Pressure corrector loop
            while (pimple.correct())
            {
                if (pimple.consistent())
                {
                    #include "pcEqn.H"
                }
                else
                {
                    #include "pEqn.H"
                }
            }

            if (pimple.turbCorr())
            {
                turbulence->correct();
            }
        }

        rho = thermo.rho();

        runTime.write();

        runTime.printExecutionTime(Info);
    }

    Info<< "End\n" << endl;

    return 0;
}


// ************************************************************************* //

\end{lstlisting}
\subsection{Discretisation}

Considering the 1D control volume, and a 1D version of the divergence 
theorem, we can write:

\begin{equation*} 
	\Delta x A_{XS} \frac{\partial \rho}{\partial t} 
	+ \Delta x \frac{\partial \dot{m}}{\partial x} = 0
\end{equation*}

\begin{equation*} 
	V \frac{\partial \rho}{\partial t} 
	+ \sum_j \dot{m}_{j, face} = 0
\end{equation*}

As with before, we must estimate $\dot{m}_{j,face}$ using the centroid 
values. We can use upwinding interpolation, which would make the most sense.

That means, using p to denote the value at the centroid of the cell,
and n to denote value at centroid of neighbouring cell.

At the face that is going to some adjacent cell,

If $\dot{m}_p \cdot \hat{S} > 0$:

$$\dot{m}_{j,face} = \dot{m}_p$$

else 


$$\dot{m}_{j,face} = \dot{m}_n$$

$\hat{S}$ is the unit area surface normal vector that points away from the 
centroid of the cell.

\subsection{Step 1: Density Equation}
For temporal time stepping:

\begin{equation*} 
	V \frac{\rho^{t + \Delta t} - \rho^t}{\Delta t} 
	+ \sum_j \dot{m}_{j, face} = 0
\end{equation*}

We have an explicit equation that allows us to determine a guess 
in density for the next iteration if we were to do it explicitly.

This is the first step in the rhoPimpleFoam algorithm. (see line 169)

If there are any flows in or out of here, we can incorporate it as 
well into this equation.

Of course, $\rho^{t + \Delta t}$ is not exactly the correct term, 
as we haven't completely solved and iterated the new density fields.
$\rho^{predicted}$ should be the correct term. So what we have now 
is an intermediate value:

\begin{equation*} 
	V \frac{\rho^{predicted} - \rho^t}{\Delta t} 
	+ \sum_j \dot{m}_{j, face} = 0
\end{equation*}

\subsection{Step 2: Momentum Predictor Step}

The momentum predictor step in OpenFOAM is as follows:
\begin{lstlisting}
	
// Solve the Momentum equation

MRF.correctBoundaryVelocity(U);

tmp<fvVectorMatrix> tUEqn
(
	fvm::ddt(rho, U) + fvm::div(phi, U)
	+ MRF.DDt(rho, U)
	+ turbulence->divDevRhoReff(U)
	==
	fvOptions(rho, U)
);
fvVectorMatrix& UEqn = tUEqn.ref();

UEqn.relax();

fvOptions.constrain(UEqn);

if (pimple.momentumPredictor())
{
	solve(UEqn == -fvc::grad(p));

	fvOptions.correct(U);
	K = 0.5*magSqr(U);
}


\end{lstlisting}
For the momentum predictor, we start with:
\begin{equation*} 
	\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Now, we know that the densities have been predicted based on mass balance,
we can do a momentum predictor step. Note that we ignore the 
$\frac{\partial \dot{m}}{\partial t}$ term just like modelica. This makes it 
easier for us. The only thing for us to discrete here is the pressure 
equation. We will use divergence theorem again by integrating each term 
by volume.

However, the 1D form has already been integrated across the areas,
so we only need do is over dx:

\begin{equation*} 
	\Delta x \frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+\rho g A_{XS} \int dx  \frac{\partial z}{\partial x} 
	+ A_{XS} \int dx  \frac{\partial P}{\partial x} = 0
\end{equation*}

\begin{equation*} 
	\Delta x \frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+\rho g A_{XS} \Delta z
	+ A_{XS} \Delta P_{x,face} = 0
\end{equation*}

For this, we can do some pressure interpolation using some weighting 
factor $\alpha$:

$$\Delta P_{x,face} = \alpha P_p + (1-\alpha) P_n $$

OpenFOAM tends to correlate $\alpha$ based on the distance from the cell 
centroid to the boundary. In the 1D case, $\alpha$ is simply based on 
the ratios of the lengths of each 1D cell.

For this, we can see 


$$ 1 - \alpha = \frac{\text{cell P length}}{
	\text{cell P length} + \text{cell N length}
} $$

Hence, 


$$ \alpha = \frac{\text{cell N length}}{
	\text{cell P length} + \text{cell N length}
} $$

Now, the density would be based on the previous density predictor step, 
but the pressures and other properties would be based on the previous 
timestep.


\begin{equation*} 
	\Delta x \frac{f_{fanning} P_w}{2 \rho A_{XS}^2} 
	\dot{m}^{predicted} |\dot{m}^{predicted}|
	+\rho g A_{XS} \Delta z
	+ A_{XS} \Delta P_{x,face} = 0
\end{equation*}

As for whether we should take the positive or negative answer of 
$|\dot{m}^{predicted}|$, that should depend on the sign of the total 
pressure change.

Now, it is evident that this is a nonlinear equation. So we want to 
linearise this equation just like for PIMPLE. We shall use the mass 
flowrates from the previous timesteps for the fanning friction factor 
and $\dot{m}$.

\begin{equation*} 
	\Delta x \frac{f_{fanning}^t P_w}{2 \rho^{predicted} A_{XS}^2} 
	\dot{m}^{t} |\dot{m}^{predicted}|
	+\rho^{predicted} g A_{XS} \Delta z
	+ A_{XS} \Delta P_{x,face}^t = 0
\end{equation*}

Now, solving for this is kind of cumbersome due to the presence of 
the modulus sign, it may be better to do:

\begin{equation*} 
	\Delta x \frac{f_{fanning}^t P_w}{2 \rho^{predicted} A_{XS}^2} 
	\dot{m}^{predicted} |\dot{m}^{t}|
	+\rho^{predicted} g A_{XS} \Delta z
	+ A_{XS} \Delta P_{x,face}^t = 0
\end{equation*}


Of course, if area and such were changing, we could use this from the last 
timestep. Nevertheless, this is a control volume, and we should not 
expect area to change. 

This makes it relatively easy to obtain $\dot{m}^{predicted}$ as the steps 
are quite explicit.

\subsection{Step 3: Energy Equation}

With the predicted mass flowrates, and densities, we can then
consider the energy equation.

\begin{equation*} 
	\rho A_{XS} \frac{\partial h_{enthalpy}}{\partial t} 
	+ A_{XS} \frac{\partial P}{\partial t} = 
	- \rho A_{XS} u \frac{\partial h_{enthalpy}}{\partial x}
	+ q_{wall} \frac{A_{wall}}{L}
\end{equation*}


\begin{equation*} 
	\rho A_{XS} \frac{\partial h_{enthalpy}}{\partial t} 
	+ A_{XS} \frac{\partial P}{\partial t} = 
	- \dot{m} \frac{\partial h_{enthalpy}}{\partial x}
	+ q_{wall} \frac{A_{wall}}{L}
\end{equation*}


Let us integrate across dx:

\begin{equation*} 
	\rho V \frac{\partial h_{enthalpy}}{\partial t} 
	+ V \frac{\partial P}{\partial t} = 
	- \sum_j \dot{m}_j h_j
	+ q_{wall} A_{wall}
\end{equation*}

Now, this is tricky because we have two time derivatives. 
One based on pressure, one based on enthalpy. What do we do with this?

Based on the OpenFOAM EEqn.H in rhoPimpleFoam:

\begin{lstlisting}
{
    volScalarField& he = thermo.he();

    fvScalarMatrix EEqn
    (
        fvm::ddt(rho, he) + fvm::div(phi, he)
      + fvc::ddt(rho, K) + fvc::div(phi, K)
      + (
            he.name() == "e"
          ? fvc::div
            (
                fvc::absolute(phi/fvc::interpolate(rho), U),
                p,
                "div(phiv,p)"
            )
          : -dpdt
        )
      - fvm::laplacian(turbulence->alphaEff(), he)
     ==
        fvOptions(rho, he)
    );

    if (MRF.active())
    {
        EEqn += fvc::div(MRF.phi(), p);
    }

    EEqn.relax();

    fvOptions.constrain(EEqn);

    EEqn.solve();

    fvOptions.correct(he);

    thermo.correct();
}
\end{lstlisting}

It appears we have to calculate a pressure gradient first.
Not sure where we get that from though!

Now, for the \textbf{first} iteration, dpdt is zero, because there is 
no prior pressure from which to calculate it.

In subsequent steps though, p will be updated. And one can have a dpdt as 

\begin{equation*}
	\frac{\partial p}{\partial t} = \frac{p^{t} - p^{t - \Delta t}}{\Delta t}
\end{equation*}

\textbf{Equation of State:}


Now, with this, we will then be able to solve for enthalpy at the next 
timestep.

Now, we have predicted a new density $\rho$, a new set of mass flowrates 
$\dot{m}$ and a new enthalpy $h$. With density and enthalpy, in theory, 
one should be able to obtain a new pressure field.

$$p = p(\rho,h)$$

For a ideal gas equation of state, this is quite easy if the heat capacity 
equation is straightforward, as pressure is explicitly 
in terms of specific volume $1/\rho$ and temperature. 

However, for cubic equation of state, then we need to find temperature 
iteratively from $h$. 

And for steam tables, we need to iteratively find the temperature which 
gives us the correct density $\rho$ using the $p,h$ equations.

Hence, for steam tables, we need to vary $p$ such that:

$$ \rho(p,h) - \rho_{set} = 0 $$

By doing so, we will obtain our pressure and (p,h) point.

Now, this step is not done in the EEqn.H step, the algorithm calls 
for finding the temperature first from the thermodynamic states of 
$\rho$ and $h$ so that we will get a $\rho,T$ point. This is ideal for 
putting into a cubic equation of state to find p.


\subsection{Step 4: pEqn}

Now, we enter the PISO part, this is the pressure corrector step.
Essentially, it is the momentum equation substituted into the pressure 
equation.

In the pEqn.H, we see pimpleFoam accounts for transonic equations, but 
we shall ignore that. We get:

\begin{lstlisting}
if (!pimple.SIMPLErho())
{
    rho = thermo.rho();
}

// Thermodynamic density needs to be updated by psi*d(p) after the
// pressure solution
const volScalarField psip0(psi*p);

volScalarField rAU(1.0/UEqn.A());
surfaceScalarField rhorAUf("rhorAUf", fvc::interpolate(rho*rAU));
volVectorField HbyA(constrainHbyA(rAU*UEqn.H(), U, p));

if (pimple.nCorrPISO() <= 1)
{
    tUEqn.clear();
}

surfaceScalarField phiHbyA
(
    "phiHbyA",
    fvc::interpolate(rho)*fvc::flux(HbyA)
  + MRF.zeroFilter(rhorAUf*fvc::ddtCorr(rho, U, phi, rhoUf))
);

fvc::makeRelative(phiHbyA, rho, U);
MRF.makeRelative(fvc::interpolate(rho), phiHbyA);

// Update the pressure BCs to ensure flux consistency
constrainPressure(p, rho, U, phiHbyA, rhorAUf, MRF);

// NOTE: I'm skipping the transonic part

{
    fvScalarMatrix pDDtEqn
    (
        fvc::ddt(rho) + psi*correction(fvm::ddt(p))
      + fvc::div(phiHbyA)
     ==
        fvOptions(psi, p, rho.name())
    );

    while (pimple.correctNonOrthogonal())
    {
        fvScalarMatrix pEqn(pDDtEqn - fvm::laplacian(rhorAUf, p));

        pEqn.solve(p.select(pimple.finalInnerIter()));

        if (pimple.finalNonOrthogonalIter())
        {
            phi = phiHbyA + pEqn.flux();
        }
    }
}


// Explicitly relax pressure for momentum corrector
p.relax();

U = HbyA - rAU*fvc::grad(p);
U.correctBoundaryConditions();
fvOptions.correct(U);
K = 0.5*magSqr(U);

if (pressureControl.limit(p))
{
    p.correctBoundaryConditions();
}

thermo.correctRho(psi*p - psip0, rhoMin, rhoMax) ;

#include "rhoEqn.H"
#include "compressibleContinuityErrs.H"

rho = thermo.rho();

// Correct rhoUf if the mesh is moving
fvc::correctRhoUf(rhoUf, rho, U, phi);

if (thermo.dpdt())
{
    dpdt = fvc::ddt(p);

    if (mesh.moving())
    {
        dpdt -= fvc::div(fvc::meshPhi(rho, U), p);
    }
}


\end{lstlisting}

We recall that the corrector step was done by substituting the expressions 
in for momentum balance into the mass conservation equations.


So we have:
\begin{equation*} 
	V \frac{\rho^{t + \Delta t} - \rho^t}{\Delta t} 
	+ \sum_j \dot{m}_{j, face} = 0
\end{equation*}

\begin{equation*} 
	\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Now, we just had intermediate guesses from 
\begin{itemize}
	\item density equation $\rho$
	\item the momentum predictor, predicting $\dot{m}$
	\item energy equation, predicting $h$
\end{itemize}

The next few steps use the pressure-velocity coupling steps from the 
SIMPLE and PISO algorithms, but there is a need to update it for density.

Recall for \textbf{incompressible} 
that the momentum equation can be written as:
\begin{equation*}
	\textbf{u}_p^{predictor} = 
	\frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
	+ \frac{\nabla \cdot \tau /\rho}{a_p}
	+ g/a_p
\end{equation*}

The pressure equation is done by substituing this equation into 
the continutity equation:

\begin{equation*}
	\nabla \cdot \mathbf{u} = 0
\end{equation*}
\begin{equation*}
	\nabla \cdot \left(
		\frac{\textbf{H}(\textbf{u}^t)}{a_p}
		- \frac{\nabla P/\rho}{a_p}
		+ \frac{\nabla \cdot \tau /\rho}{a_p}
		+ g/a_p
	\right) = 0
\end{equation*}

By doing so, we can have guesses for the pressure in terms of momentum,
or velocity guesses:

\begin{equation*}
	\nabla \cdot \frac{\nabla P/\rho}{a_p} 
	=
	\nabla \cdot \left(
		\frac{\textbf{H}(\textbf{u}^t)}{a_p}
		+ \frac{\nabla \cdot \tau /\rho}{a_p}
		+ g/a_p
	\right) 
\end{equation*}

This was the original equation mentioned in Jasak's thesis 
\citep{jasak1996error}.

That is well and good, but now we have another layer of complication.
We move from a pressure based solver to a density based solver.
Bone outlines this thought process well \citep{bone2020comparative}.
Bone then cites Kurganov et al. for coming up with the 
central schemes (rhoCentralFoam) \citep{kurganov2000new}.

Bone mentiones rhoSimpleFoam on page 6-7 \citep{bone2020comparative}, 
this is essential for understanding rhoPimpleFoam. Now, in page 3 of 
the thesis, Bone mentions about steady state compressible equations.

In that regard, continutity equation is similar to SIMPLE algorithms
for incompressible flow, except that $\rho$ is added in. Time derivative 
terms are not accounted for.

For page 7 onwards, Bone writes about the PISO algorithm for compressible 
flow \citep{bone2020comparative}. This is more relevant to our case.
However, Bone does not elaborate on how density is solved for in 
rhoPimpleFoam. This is also not covered in the rhoCentralFoam part of the 
thesis. However, the solution procedure is standard, density, momentum,
energy, then EOS \citep{bone2020comparative}. 

Clues from sonicFoam can also be obtained from Canteros 
\citep{canteros2024review} as sonicFoam uses the PIMPLE algorithm like 
rhoPimpleFoam.


Now, anyway, the first thing that is done is to use the Equation of 
state to find a pressure.

So now, we have these values which are guessed from before:

\begin{itemize}
	\item From density eqn: $\rho^{predicted}$
	\item From momentum predictor: $\dot{m}^{predicted}$
	\item From energy equation: $h^{predicted}$
	\item From Equation of State: $\rho,T$ from $\rho, h$
\end{itemize}

Now we have:

\begin{itemize}
	\item $p$ from $\rho, T$ (or $\rho,h$, should not be an issue either)
\end{itemize}


So far, there is minimal coupling between momentum and mass balance 
equation. Hence, we are using these two equations to form the pressure 
equation from before:

\begin{equation*} 
	V \frac{\rho}{\Delta t} 
	+ \sum_j \dot{m}_{j, face} = 0
\end{equation*}

The face fluxes will be dealt with using the momentum predictor as before.
It is the pressure time derivative term that is slightly more tricky. 
Now, to deal with this, the equation of state can be brought in. If we 
deal with some generic equation of state with a compressibility 
factor $z$:

\begin{equation*}
	\rho = \rho(p,T) = \frac{p}{z RT}
\end{equation*}

The term $\frac{1}{zRT}$ is known as $\psi$ in pEqn.H:

\begin{equation*}
	\rho = \psi p
\end{equation*}

Suppose now that we can form:


\begin{equation*}
	d \rho = \psi dp + p d\psi
\end{equation*}

The time derivative becomes equation \ref{eqn:time-derivative-of-density}:

\begin{equation} \label{eqn:time-derivative-of-density}
	\frac{\partial \rho}{\partial t} = \psi \frac{\partial p}{\partial t} 
	+ p \frac{\partial \psi}{\partial t}
\end{equation}

Now, as far as we are concerned, we've already dealt with the energy, and 
the thermodynamic changes do not need to change and such.

Now, in the first parts of the predictor step, we have already accounted 
for changes in density due to energy (temperature) increase at constant 
pressure. This is the second term. The $p$ there will be the pressure 
at the previous timestep.

Now, due to temperature increase, we already accounted for the change 
by changing density due to temperature

\begin{equation*}
	\frac{\partial \rho}{\partial t} = \psi \frac{\partial p}{\partial t} 
	+ p \frac{\psi^{\text{after EEqn.H}} - \psi^\text{previous timestep}}{\Delta t}
\end{equation*}

Now, if we were to convert this back to expressions of density:

\begin{equation*}
	\frac{\partial \rho}{\partial t} = \psi \frac{\partial p}{\partial t} 
	+  \frac{\rho^{\text{after EEqn.H}} - \rho^\text{previous timestep}}{\Delta t}
\end{equation*}

It would be evident now, based on the thermodynamic derivatives why 
there are two components of the density change, one due to increased 
enthalpy, one due to pressure.


\begin{equation*}
	\frac{\partial \rho}{\partial t} = \psi 
	\frac{p^\text{next timestep} - p^\text{previous timestep}}{\Delta t} 
	+ \frac{\rho^{\text{after EEqn.H}} - \rho^\text{previous timestep}}{\Delta t}
\end{equation*}

From this regard, the equations now make physical sense. And it is more 
evident why there are $\rho$ present in the OpenFOAM rhoPimpleFoam 
equations. Thus, the pressure equation combines the mass balance,
equation of state, and momentum balance, in order to predict the pressure 
field at the next timestep.

Note that after the pressure equation is solved, 
the mass conservation equation in the density equation is solved again, 
and then the density is updated using the equation of state.

Now, iterating over these equations many times makes guessing 
$\rho^{\text{after EEqn.H}}$ much more accurate and consistent with 
the pressure changes to be done. This results in  
equation \ref{eqn:time-derivative-of-density-discretised}:

\begin{equation} \label{eqn:time-derivative-of-density-discretised}
	\frac{\partial \rho}{\partial t} = \psi 
	\frac{p^\text{n+1} - p^\text{n}}{\Delta t} 
	+ \frac{\rho^{\text{after EEqn.H}} - \rho^\text{n}}{\Delta t}
\end{equation}

With this, we get better estimates for pressure at the next timestep.

\subsection{relaxation factors}

Note that there are explicit relaxation factors applied for pressure 
in the pEqn.H.



\subsection{Translation for rhoPimpleFoam to 1D case}

With the background of rhoPimpleFoam.C in mind, let us now consider how 
we can properly discretise our equations obtained from Modelica.

Let's start with pEqn.H, convert that to a rust equivalent for 1D 
flows, and work our way out.
\subsubsection{pEqn equivalent}

We know that the momentum predictor equation in 3D is as follows:

\begin{equation*}
	\textbf{u}_p^{predictor} = \frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
\end{equation*}

Let's develop a momentum predictor equation. From modelica in 
equation \ref{eqn:modelica-momentum-balance-1D} \citep{Casella2003} 
we have:

\begin{equation*} 
	\frac{\partial \dot{m}}{\partial t}
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

For this, I'd rather have momentum predictor equations in terms 
of mass balance. Let's start with the time dependent terms:

\begin{equation*} 
	\frac{\dot{m}^{t + \Delta t} - \dot{m}^t}{\Delta t}
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Of course, in the pEqn, we don't actually predict the next timestep 
mass flow rate straight away. We actually predict the properties at the 
next iteration. This we denote as $n$, and $n+1$.


\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Now, we can move on to the next terms. We notice the momentum advection 
terms are visibly missing compared to rhoPimpleFoam.C. This is indeed a 
simplification. Will momentum advection terms be significant? 
As I googled, AI reported that this is significant in high speed flows. (No 
citation here, may add later). I would like to add it back to keep things 
consistent with rhoPimpleFoam.C.

How are we doing it? Let's start with the substantial derivative:

\begin{equation*}
	\rho \frac{D \mathbf{u}}{D t} 
	= 
	\rho \frac{\partial \mathbf{u}}{\partial t}
	+ 
	\rho (\mathbf{u} \cdot \nabla) \mathbf{u}
\end{equation*}

Let's change this to 1D:

\begin{equation*}
	\rho \frac{D u}{D t} 
	= 
	\rho \frac{\partial u}{\partial t}
	+ 
	\rho u \frac{\partial}{\partial x} u
\end{equation*}

Multiply cross sectional area to both sides:
\begin{equation*}
	\rho A_{XS} \frac{D u}{D t} 
	= 
	\rho A_{XS} \frac{\partial u}{\partial t}
	+ 
	\rho A_{XS} u \frac{\partial}{\partial x} u
\end{equation*}

\begin{equation*}
	 \frac{D \dot{m}}{D t} 
	= 
	\frac{\partial \dot{m}}{\partial t}
	+ 
	u \frac{\partial \dot{m}}{\partial x} 
\end{equation*}


\begin{equation*}
	 \frac{D \dot{m}}{D t} 
	= 
	\frac{\partial \dot{m}}{\partial t}
	+ 
	\frac{\dot{m}}{A_{XS} \rho} \frac{\partial \dot{m}}{\partial x} 
\end{equation*}


Hence, adding this back:


\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}}{A_{XS} \rho} \frac{\partial \dot{m}}{\partial x} 
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Now, for the advection term, it is evident that some discretisation 
is required because we need to find the mass flowrates at the faces 
using the divergence theorem. 


\begin{equation*}
	\iiint \nabla \cdot \mathbf{u} dV = \iint \mathbf{u} \cdot d \vec{S}
\end{equation*}

Of course, if we were to include density:

\begin{equation*}
	\rho \iiint \nabla \cdot \mathbf{u} dV = \rho \iint \mathbf{u} \cdot d \vec{S}
\end{equation*}

Now, in a 1D case:

\begin{equation*}
	\rho A_{XS} \int \frac{\partial}{\partial x} u dx = \rho \iint u \cdot d \vec{S}
\end{equation*}

If we were to replace with sums:


\begin{equation*}
	\rho A_{XS} \int \frac{\partial}{\partial x} u dx 
	= \rho \sum_j u_j \cdot \vec{S}_j
\end{equation*}


\begin{equation*}
	\int \frac{\partial \dot{m}}{\partial x}  dx 
	= \rho \sum_j u_j \cdot \vec{S}_j
\end{equation*}

If we were to derive an expression back for 
$\frac{\partial \dot{m}}{\partial x}$:

\begin{equation*}
	\int \frac{\partial \dot{m}}{\partial x}  dx 
	= \rho \sum_j u_j \cdot \vec{S}_j
\end{equation*}

Differentiating both sides with respect to $x$ and noting a consistent 
cross sectional area across the control volume:

\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{\partial }{\partial x} \rho \sum_j u_j \cdot \vec{S}_j
\end{equation*}


\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{1}{A_{XS}} \frac{\partial }{\partial x} \sum_j \dot{m}_j \cdot \vec{S}_j
\end{equation*}

Note that the dot product implies a sign convention. For divergence, 
it measures the rate at which a flux diverges (spreads out). In this case,
it means outflow of flux (or mass) from a system. Hence $\vec{S}$ necessarily 
points outwards from the centroid of the cell.

Now, for mass flowrates at the faces, $\dot{m}$, we need some sort 
of interpolation scheme to have it in terms of the mass flowrates at 
the centroids. For velocity related interpolation schemes, upwind is used 
for stability but suffers from diffusivity issues 
\citep{alves2003convergent}.

Let's consider this:

\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{1}{A_{XS} \Delta x}  \sum_j \dot{m}_j \cdot \vec{S}_j
\end{equation*}

Let's do upwind interpolation and assume $A_{XS}$

\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{1}{A_{XS} \Delta x}  \sum_j \dot{m}_j \cdot \vec{S}_j
\end{equation*}


\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{1}{A_{XS} \Delta x}  \sum_j \dot{m}_j \cdot \vec{S}_j
\end{equation*}

Note: $m_j$ is flow OUT of the control volume:


\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{\sum_j \dot{m}_{j,out}}{\Delta x}  
\end{equation*}
In the end, we simplify to this.

If we take mass flowrate as positive in x direction:

\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{\dot{m}_{x+0.5 \Delta x} - \dot{m}_{x-0.5 \Delta x}}{\Delta x}  
\end{equation*}

If we use linear interpolation:


\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{0.5*(\dot{m}_{CV+1}+\dot{m}_{CV}) - 0.5(\dot{m}_{CV}+\dot{m}_{CV-1})}{\Delta x}  
\end{equation*}

\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{0.5\dot{m}_{CV+1} - 0.5\dot{m}_{CV-1}}{\Delta x}  
\end{equation*}

Basically, the CV+1 demarcates the properties in the neigbouring cell in 
the positive $x$ direction, whereas CV-1 demarcates properties in the 
neigbouring cell in the negative $x$ direction.

This provides a linear way to estimate the mass flowrates at the faces, 
and is essentially the same as the central difference approximation.

Upwinding means that the mass flowrate depends on the flow direction.


Meaning to say, if there is influx into the cell, the mass flowrate at 
the face is the mass flowrate at the neighbouring cell centroid. If 
there is flow out of the cell, the mass flowrate at the face is the mass 
flowrate at cell centroid.

If flow is positive in the x direction:
\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{\dot{m}_{CV} - \dot{m}_{CV-1}}{\Delta x}  
\end{equation*}

But if flow is in the negative x direction:
\begin{equation*}
	\frac{\partial \dot{m}}{\partial x}  
	= \frac{\dot{m}_{CV+1} - \dot{m}_{CV}}{\Delta x}  
\end{equation*}

Upwind schemes have to depend on this if else statement in the positive 
and negative x direction. This is essentially the forward and backward 
difference approximation. It is first order in nature.

From a programming point of view, the upwind scheme is more slightly more 
difficult to program as compared to the linear approximation.

So we can add this in:

\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}}{A_{XS} \rho} \frac{\sum_j \dot{m}_j}{\Delta x} 
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Where $\sum_j \dot{m}_j$ is sum of mass flowrates out of the cell.
In terms of time or iteration discretisation, we can linearise it

\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}^n}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n+1}}{\Delta x} 
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

For the friction term, we note that the fanning friction factor depends 
on $\dot{m}$. This will be based on $\dot{m}$ at the previous iteration.


\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}^n}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n+1}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} \dot{m}^{n+1} |\dot{m}^n|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

The rest of the terms are in the previous iteration. Pressure should also 
use a Gauss linear approximation.


\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}^n}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n+1}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} \dot{m}^{n+1} |\dot{m}^n|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} = 0
\end{equation*}


Hence, our momentum predictor equation becomes:
\begin{equation*} 
	\frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}^n}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n+1}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} \dot{m}^{n+1} |\dot{m}^n|
	=
	- \rho g A_{XS} \frac{\partial z}{\partial x}
	- A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} 
\end{equation*}

Now, we note that $\dot{m}_j$ necessarily implies that mass flowrates 
of neighbouring cells are considered. Thus, solution of this system 
of equations will become semi implicit as there is a matrix solution to 
this.

To simplify this, we will do the following:

\begin{align*}
	& \frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}^{n+1}}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} \dot{m}^{n+1} |\dot{m}^n|  \\ 
	&
	=
	- \rho g A_{XS} \frac{\partial z}{\partial x}
	- A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} \\
\end{align*}


In doing so, all mass flowrates can be guessed explicitly. Of course, 
we can do Gauss-Siedel iterations to update the guesses of the mass flowrates.
We can do so until the mass flowrates reached some sort of equilibrium.
Otherwise, we can just let the iterations continue till some 
convergence is reached.

Now, for our momentum predictor we start by going back to the momentum 
predictor in OpenFOAM. We note that the momentum predictor equation is:
\begin{equation*}
	\textbf{u}_p^{predictor} = \frac{\textbf{H}(\textbf{u}^t)}{a_p}
	- \frac{\nabla P/\rho}{a_p}
\end{equation*}

So, we need to arrange things like that of the momentum predictor.
This means we move the explicit constants to the right hand side:

\begin{align*}
	& \frac{\dot{m}^{n+1} - \dot{m}^n}{\Delta t}
	+ \frac{\dot{m}^{n+1}}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} \dot{m}^{n+1} |\dot{m}^n|  \\ 
	&
	=
	- \rho g A_{XS} \frac{\partial z}{\partial x}
	- A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} \\
\end{align*}


\begin{align*}
	& 
	\frac{\dot{m}^{n+1} }{\Delta t}
	+ \frac{\dot{m}^{n+1}}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} \dot{m}^{n+1} |\dot{m}^n|  \\ 
	&
	=
	\frac{ \dot{m}^n}{\Delta t}
	- \rho g A_{XS} \frac{\partial z}{\partial x}
	- A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} 
	\\
\end{align*}

\begin{align*}
	& 
	\dot{m}^{n+1}
	\left(
		\frac{1 }{\Delta t}
		+ \frac{1}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n}}{\Delta x} 
		+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} |\dot{m}^n|  
	\right)
	\\ 
	&
	=
	\frac{ \dot{m}^n}{\Delta t}
	- \rho g A_{XS} \frac{\partial z}{\partial x}
	- A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} 
	\\
\end{align*}


It is clear from this that H is:

\begin{equation*}
	H = \frac{ \dot{m}^n}{\Delta t}
	- \rho g A_{XS} \frac{\partial z}{\partial x}
\end{equation*}

And the equivalent of $\nabla P/\rho$ is:


\begin{equation*}
	A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} 
\end{equation*}

Then, $a_P$ is:

\begin{equation*}
	a_P = \frac{1 }{\Delta t}
	+ \frac{1}{A_{XS} \rho} \frac{\sum_j \dot{m}_j^{n}}{\Delta x} 
	+\frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} |\dot{m}^n|  
\end{equation*}

Now, the momentum predictor is:

\begin{equation} \label{eqn:momentum-predictor-1d-mass-flowrate}
	\dot{m}_p^{n+1} = \frac{H(\dot{m}^n)}{a_p} + 
	\frac{1}{a_p}
	A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} 
\end{equation}

Now, as with Modelica, we should have the option to change the time 
derivative to zero as well as the momentum advection term using some if 
else statement. In this case, $a_P$ is simplified to:

\begin{equation*}
	a_P = \frac{f_{fanning}^n P_w}{2 \rho A_{XS}^2} |\dot{m}^n|  
\end{equation*}

And $H$ is simplified to:
\begin{equation*}
	H = - \rho g A_{XS} \frac{\partial z}{\partial x}
\end{equation*}

With this, it should be sufficient to obtain a momentum predictor equation,
which takes in all the parameters, then spits out the mass flowrates 
at the next iteration.


Now, in the case of continuity, we need to 
know $\nabla \cdot \mathbf{u} = 0$.

However, this is definitely not the case in here. We can start from 
a 3D density equation, in order to arrive at the pressure coupling equation 
pEqn.H.

\begin{equation*}
	\frac{\partial \dot{m}}{\partial t} + \sum_j \dot{m}_j = 0
\end{equation*}

Using the same convention, $\sum_j \dot{m}_j$ is the mass flowrate out 
of the differential control volume. Since volume is constant:

\begin{equation*}
	V \frac{\partial \rho}{\partial t} + \sum_j \dot{m}_j = 0
\end{equation*}

\begin{equation*}
	A_{XS} \Delta x  \frac{\partial \rho}{\partial t} + \sum_j \dot{m}_j = 0
\end{equation*}
\begin{equation*}
	A_{XS} \frac{\partial \rho}{\partial t} + \frac{1}{\Delta x} \sum_j \dot{m}_j = 0
\end{equation*}

Or if we don't discretise it:

\begin{equation*}
	A_{XS} \frac{\partial \rho}{\partial t} + \frac{\partial \dot{m}}{\partial x} = 0
\end{equation*}
This is exactly the Modelica equations.

Now, we need an expression for $\frac{\partial \dot{m}_j}{\partial x}$. The 
only way this is possible is from the momentum advection term. Otherwise, 
it is very hard to have pressure velocity coupling. Note that the point 
of the pEqn.H is to predict pressures.


\begin{equation*} 
	\frac{\partial \dot{m}}{\partial t}
	+ \frac{\dot{m}}{A_{XS} \rho} \frac{\partial \dot{m}}{\partial x} 
	+\frac{f_{fanning} P_w}{2 \rho A_{XS}^2} \dot{m} |\dot{m}|
	+ \rho g A_{XS} \frac{\partial z}{\partial x}
	+ A_{XS} \frac{\partial P}{\partial x} = 0
\end{equation*}

Now, this isn't going to cut it. The better way, just like in the 
rhoPimpleFoam equations, is to just substitute 
the momentum predictor equation straight in in 
equation \ref{eqn:momentum-predictor-1d-mass-flowrate}:

\begin{equation*} 
	\dot{m}_p^{n+1} = 
	\frac{H(\dot{m}^n)}{a_p} + 
	\frac{1}{a_p}
	A_{XS} \frac{0.5 P^{n}_{CV+1}-0.5P^n_{CV-1}}{\Delta x} 
\end{equation*}

Now, it is noteworthy here that we are trying to predict the pressure 
at the next iteration. So we need to change the notation slightly.

\begin{equation*} 
	\dot{m}_p^{n+1} = 
	\frac{H(\dot{m}^n)}{a_p} + 
	\frac{1}{a_p}
	A_{XS} \frac{0.5 P^{n+1}_{CV+1}-0.5P^{n+1}_{CV-1}}{\Delta x} 
\end{equation*}

We can even go as far to say that the mass flowrate we use for H is 
based on the momentum predictor:
\begin{equation*} 
	\dot{m}_p^{n+1} = 
	\frac{H(\dot{m}^{predictor})}{a_p} + 
	\frac{1}{a_p}
	A_{XS} \frac{0.5 P^{n+1}_{CV+1}-0.5P^{n+1}_{CV-1}}{\Delta x} 
\end{equation*}

Let's substitute this in the mass balance:
\begin{equation*}
	A_{XS} \frac{\partial \rho}{\partial t} 
	+ \frac{\partial \dot{m}}{\partial x} = 0
\end{equation*}

We specifically note that the mass flowrate $\dot{m}$ must be 
at the next iteration.
\begin{equation*}
	A_{XS} \frac{\partial \rho}{\partial t} 
	+ \frac{\partial \dot{m}^{n+1}}{\partial x} = 0
\end{equation*}

To yield:
\begin{equation*}
	A_{XS} \frac{\partial \rho}{\partial t} 
	+ \frac{ 
		\partial 
		\left(
			\frac{H(\dot{m}^{predictor})}{a_p} + 
			\frac{1}{a_p}
			A_{XS} \frac{0.5 P^{n+1}_{CV+1}-0.5P^{n+1}_{CV-1}}{\Delta x} 
		\right)
	}{\partial x} = 0
\end{equation*}

We note that these properties must be evaluated at the faces. 
Now, the H terms at the faces 
can be evaluated using the upwind schemes, as well as 
the $a_P$ terms. This will help in interpolating the terms at the faces.
For pressure, we can use a Gauss linear scheme which is second order 
accurate.

Now, we still must consider the time derivative of density.

We note that density is a function of both temperature and pressure.
Recall equation \ref{eqn:time-derivative-of-density}:

\begin{equation*} 
	\frac{\partial \rho}{\partial t} = \psi \frac{\partial p}{\partial t} 
	+ p \frac{\partial \psi}{\partial t}
\end{equation*}

Recall, that we discretise this to equation
\ref{eqn:time-derivative-of-density-discretised}:

\begin{equation*} 
	\frac{\partial \rho}{\partial t} = \psi 
	\frac{p^\text{n+1} - p^\text{n}}{\Delta t} 
	+ \frac{\rho^{\text{after EEqn.H} - \rho^\text{n}}}{\Delta t}
\end{equation*}

%\begin{verbatim}
%From AI:
%pEqn.H Execution Flow:
%━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
%
%1. Update ρ from thermo (ρ = ψp)
%2. Store ψp₀ for later density correction
%3. Compute rAU = 1/aₚ from momentum equation
%4. Construct H/A (velocity predictor without pressure)
%5. Compute predicted flux φ_HbyA = ρ(H/A)·S
%6. Apply boundary constraints
%
%7. SOLVE PRESSURE EQUATION:
%   ┌─ Assemble: ∂ρ/∂t + ψ∂p/∂t + ∇·φ_HbyA - ∇·(ρ/aₚ ∇p) = 0
%   │
%   └─ Non-orthogonal loop:
%       ├─ Solve for p (implicit Laplacian)
%       └─ Update φ = φ_HbyA - (ρ/aₚ)(∇p)
%
%8. Relax pressure: p = αₚp_new + (1-αₚ)p_old
%9. Correct velocity: U = H/A - (1/aₚ)∇p
%10. Apply pressure limits
%11. Correct density: ρ += ψ(p_new - p_old)
%12. Check continuity errors
%13. Final ρ = thermo.rho() update
%14. **Calculate dpdt = (p_new - p_old)/Δt** ← For next EEqn.
%\end{verbatim}

\printbibliography

% \appendix
\end{document}
